---
title: "KZ_HW_001"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown Exploration!!

Hello! Welcome to my first homework assignment. In this assignment I will be exploring the **gapminder** dataset (wahoo) 

Before we begin... if you have any questions about the functions you can use the help funciton. For example if we are wondering what the **install.packages** function does, we add a question mark to the front of the function and run it through the console. Here is an example:

```{r}
?install.packages
```

Our first step is to install the gapminder package using the **install.packages("gapminder")** function. I will not include it below as this is a function that you only want to run in your console once, rather than multiple times in your markdown document. You may also want to run an install for the **tidyverse** package in your console. It is a set of packages that may make your life easier in the future if you decide to make your analysis more complex

The next thing we do is load the gapminder dataset. To accomplish this we use the **library** function. We will also load the **tidyverse** package that we just downloaded

```{r}
library(gapminder)
library(tidyverse)
```

To get a summary of the dataset we use the **summary** function

```{r}
summary(gapminder)
```

If you only want a preview of the first bit of the data set you can use the **head** function

```{r}
head(gapminder)
```

The **ncol** function tells you how many columns are present while the **nrow** function tells you how many rows are present in your dataset

```{r}
ncol(gapminder)
nrow(gapminder)
```

Now we will dig deeper. 

If you only want to see a certain part of the dataset you can use the **select** function. For example if you only want to see the life expectancy per country you will do this

```{r}
select(gapminder, country, lifeExp)
```

If we decide that we want to rename a variable in the dataset (say there is a mistake or a repeat) we can use the **rename** function

```{r}
rename(gapminder, AgeOfDeath=lifeExp)
```

That is a little dark but we will go with it

If we want to order the dataset in a certain way we can use the **arrange** function. This will order the data by population

```{r}
arrange(gapminder, pop)
```

If you want this to decrease rather than increase, we use the same function but we make a small tweak

```{r}
arrange(gapminder, desc(pop))
```

A big perk of R Markdown is the the ease of statistical applications, I find it quite handy for greating plots. for example I can create a **plot** with R Markdown will spit out a visualization of it instantly

```{r}
plot(gapminder)
```


If we want to get REALLY crazy we could do something like this

```{r}
plot(gapminder$lifeExp~gapminder$gdpPercap, cex=0.5, pch=16, col=rgb(0,0,0,0.25))
```

Thanks for checking out my data exploration. These are the basics and are essential when starting with any data set 


